On the minimum leaf conductance

### 2017-08-01 manuscript.Rmd
some writing; references added

\
Notes
gmin can increase with ozone and or aerosol particles (no ref yet) - potential for interaction with air pollution. Kerstiens?

Hoad et al; strong positive relationship between g0 and RWC - is our segmented linear method hiding changes in gmin over time?

Gardingen & Grace, gcut depends on VPD (and some useful discussion on difference isolated cuticle vs. intact leaf on non-stomatal side)

@@

hygen midgaard 1954- no effect of VPD on gmin (eventual decline in RWC is very linear and does not differ with various VPD treatments). "Experiments on the effect of darkening indicated that the stomata may remain partly open even during prolonged dark periods"

\
Stomata exert a strong control on the transpiration rate of leaves. It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low. To avoid dessiccation and ultimate death, stomata can also be expected to close during periods of severe water stress. Plants cannot, however, completely halt water loss even when stomata are closed. In this review, we discuss various aspects of the minimum leaf conductance reached when stomata are essentially closed. We pay special attention to the role that minimum conductance plays in land-surface models, in particular as it forms part of the much-used 'Ball-Berry'-type class of stomatal conductance models.

@@

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. Based on this, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is the 'cuticular conductance` [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996]. Another measurement of the minimum conductance arises from leaf drying curve: leaves are picked, and gradually dried out and weighed to determine the water loss rate [@hygen1951]. This estimate of minimum conductance (which we refer to as g~min~) probably most closely reflects the dessiccation dynamics of severely droughted plants [@blackman2016], but has rarely been mentioned in the context of the Ball-Berry model.

\
In this review, we aim to compare values of the various estimates of minimum conductance, discuss components and mechanisms of minimum conductance. We place additional emphasis on sources of variation in g~min~, and the role that g~min~ can play in models of water use. Before we discuss the various definitions and measurements of minimum conductance, we first review the modelling literature on its use of g~0~ in Eq. 1 (and related approaches), and demonstrate why g~0~ needs to be included in models.

@@

\
A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to optimize the rate of photosynthesis per unit water used [@cowan1977]. The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature). Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e. spend water) when there is no photosynthetic gain. Early work on optimal stomatal conductance models completely ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999], whereas more recent work derives the optimal g~s~ and then adds on a g~0~ [@medlyn2011].

\
In the following we discuss three reasons to include g~0~ in models: the response of C~i~ at low PPFD, the response of transpiration to high leaf temperature (T~leaf~), and the dessiccation of leaves in severe drought.

\
As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD). When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

@@

\
The third reason for inclusion of a g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants dessiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000]. We return to measurements of g~min~ from leaf drying curves in Section XXX.

# Comparison of values for various definitions

@@

| Variable   |  Definition |
| ---------- | ----------- |
| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
| g~min~     | Minimum conductance measured from a leaf dry-down curve |
| g~dark~    | Night-time conductance                                  |
| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |

\
We identify four distinct definitions of minimum conductance, each of which is measured with its own set of methods. Before discussing each of the minimum conductances in detail in the following sections, we here compile literature values to compare the relative magnitude of each of the conductances.

\
Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) and of minimum conductance from leaf dry-down curves (g~min~) were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ re-calculated). We excluded measurements of g~min~ where stomatal leaf sides were sealed, because few data were available and the comparison is not consistent with individual studies where the role of stomata in g~min~ is studied (see SectionXXX). A compilation of night-time (and 'dark') conductance (g~dark~) values was taken from @lombardozzi2017. We also estimated g~0~ from Eq. 1 using the stomatal conductance model of @medlyn2011, using non-linear regression. These estimates were combined with the compilation of g~0~ values from @miner2017, who reported on xxx values from a literature compilation (each of those values was estimated with linear regression of the Ball-Berry model). Finally, we also estimated minimum values of g~s~ attained from a large database of leaf gas exchange data [@lin2015], by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as it represents actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~0~ and g~dark~, all values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011].

@@

The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain. It is clear from our literature review, g~cut~ is very much lower than other definitions of minimum conductance - it is on average twice as low as g~min~. We admit some caution in this conclusion, because g~cut~ is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996]. We briefly review a few important aspects of cuticular conductance, and to what extent we expect it to vary with species and environments. We refer to previous in-depth reviews for a more detailed treatment of other aspects of plant cuticles, such as chemical composition and morphology, and modes of water transport across the cuticle [@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001]. We return to the question on what component of g~min~ is due to g~cut~ in the next section.

Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle. The main barrier to diffusion is actually a very thin layer at the leaf surface. Because most of the resistance is located in such a thin layer, g~cut~ does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002]. One may ask why there is so much variation in cuticle thickness between plant species, if it does not affect g~cut~. The answer may lie in mechanical strength rather than water relations, as @onoda2012 reported that thicker cuticles are more resistant to tearing (he force to tear and tensile strength both approx proportional to cuticle thickness).

@@

- @riederer2001 @schuster2016 correlation between temperature and gcut (non-stomatal cuticles)

@@

Sack et al. 2010 protocol for measurement

- similar to minimum gs observed for the Lin dataset
- plastic trait

@@

- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59). T range much wider than Blackman though... But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
page 90 few gmin values from drying curves that we can add.

>>>>

\

@@

@hygen1954 - no effect of VPD on gmin (eventual decline in RWC is very linear and does not differ with various VPD treatments). "Experiments on the effect of darkening indicated that the stomata may remain partly open even during prolonged dark periods"

@mackova2013 adding ABA to Lepidium sativum increased the chain length of cuticular waxes (but not total amounts).

@rosasanderson2014, (2013, thesis), lower gmin in drought-stressed peanut

\
Stomata exert a strong control on the transpiration rate of leaves. It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface. In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low. To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress. Plants cannot, however, completely halt water loss even when stomata are closed. In this review, we discuss various aspects of the minimum leaf conductance reached when stomata are essentially closed. We pay special attention to the role that minimum conductance plays in land-surface models, in particular as it forms part of the much-used 'Ball-Berry'-type class of stomatal conductance models.

@@

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. Based on this, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is the 'cuticular conductance` [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996]. Another measurement of the minimum conductance arises from leaf drying curve: leaves are picked, and gradually dried out and weighed to determine the water loss rate [@hygen1951]. This estimate of minimum conductance (which we refer to as g~min~) probably most closely reflects the desiccation dynamics of severely droughted plants [@blackman2016], but has rarely been mentioned in the context of the Ball-Berry model.

\
Here we review the various options to estimate g~0~ for use in models of plant water use and carbon uptake. We place an emphasis on sources of variation in g~min~, and the role that g~min~ can play in models of water use. Before we discuss the various definitions and measurements of minimum conductance, we first review the modelling literature on its use of g~0~ in Eq. 1 (and related approaches), and demonstrate why g~0~ needs to be included in models.

@@

\
A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977]. The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature). Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e. spend water) when there is no photosynthetic gain. Indeed, early work on optimal stomatal conductance models ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999]. More recent work derives the optimal g~s~, and simply adds a g~0~ to the solution [@medlyn2011].

\
In the following we discuss three reasons to include g~0~ in models: the response of C~i~ at low PPFD, the response of transpiration to high leaf temperature (T~leaf~), and the desiccation of leaves in severe drought. As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD). When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~. To see this point, we can rearrange Eq. 1 to give,

@@

\
The third reason for inclusion of a g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].

# Comparison of various definitions of minimum conductance

@@

| Variable   |  Definition |
| ---------- | ----------- |
| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
| g~min~     | Minimum conductance measured from weight loss of detached leaves |
| g~dark~    | Night-time conductance                                  |
| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |

\
Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilation by @kerstiens1996. Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from XXX papers, see Appendix. A compilation of night-time (and 'dark') conductance (g~dark~) values was taken from @lombardozzi2017. We also estimated g~0~ for each study included in a large database of leaf gas exchange data [@lin2015], by fitting Eq. 1 using the stomatal conductance model of @medlyn2011. These estimates were combined with the compilation of g~0~ values from @miner2017, who reported on xxx values from a literature compilation (each of those values was estimated with linear regression of the Ball-Berry model). Finally, we also estimated minimum values of g~s~ attained from the Lin et al. database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

\
It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~0~ and g~dark~, all values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011], and taxonomic family and order with the taxize R package [@chamberlain2013].

@@

The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain. It is clear from our literature review, g~cut~ is very much lower than other definitions of minimum conductance - it is on average twice as low as g~min~. We admit some caution in this conclusion, because g~cut~ is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996]. We briefly review a few important aspects of cuticular conductance, and to what extent we expect it to vary with species and environments. We refer to previous in-depth reviews for a more detailed treatment of other aspects of plant cuticles, such as chemical composition and morphology, and modes of water transport across the cuticle [@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001; @schuster2016b].

Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle. The main barrier to diffusion is actually a very thin layer at the leaf surface. Because most of the resistance is located in such a thin layer, g~cut~ does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002]. One may ask why there is so much variation in cuticle thickness between plant species, if it does not affect g~cut~. The answer may lie in mechanical strength rather than water relations, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).

Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in cuticular conductance (gcutisol). Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.

@riederer2001 @schuster2016b correlation between temperature and gcut (non-stomatal cuticles)

@@

@@

@@

- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59). T range much wider than Blackman though... But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
